### 2020-06-10 report/sections/introduction/motivation.tex
Report: complete introduction

Modern interactive web applications aim to provide a highly responsive user experience by minimising the communication latency between clients and servers. 
Whilst the HTTP request-response model is sufficient for retrieving static assets, applying the same stateless communication approach for interactive use cases (such as multiplayer games) introduces undesirable performance overhead from having to frequently set up new connections for client-server interactions. 
Developers have since adopted other communication transport abstractions over HTTP connections such as the WebSockets protocol \cite{WebSocketRFC} to enjoy low-latency full-duplex client-server communication in their applications over a single persistent connection. 
Enabling more complex communication patterns caters for more interactive use cases, but introduces additional concerns to the developer with respect to implementation correctness.

@@

Whilst WebSockets make this web-based implementation possible, it introduces the developer to a new family of communication errors, even for this simple game. In addition to the usual testing for game logic correctness, the developer needs to test against \textit{deadlocks} (e.g. both players waiting for each other to make a move at the same time) and \textit{communication mismatches} (e.g. player 1 sending a boolean to the game server instead of the board coordinates). The complexity of these errors, which correlate to the complexity of tests required against these errors, scale with the complexity of the communication patterns involved.

@@

We see the application of the MPST methodology to verifying interactive web applications written in TypeScript to be an interesting design space - to what extent can the MPST methodology be applied such that we leverage TypeScript language features to maximise static safety guarantees whilst preserving a flexible, natural and idiomatic workflow for TypeScript developers? Such a workflow would ultimately decrease the overhead for incorporating MPST into mainstream web development, which reduces development time by programmatically verifying implementations for communication correctness.

>>>>

Modern interactive web applications aim to 
provide a highly responsive user experience by 
minimising the communication latency between clients and servers. 
Whilst the \textit{HTTP} request-response model is 
sufficient for retrieving static assets, applying the same 
stateless communication approach for interactive use cases 
(such as multiplayer games) introduces undesirable 
performance overhead from having to frequently set up 
new connections for client-server interactions. 
Developers have since adopted other communication 
transport abstractions over HTTP connections such as the WebSockets protocol \cite{WebSocketRFC} to enjoy low-latency full-duplex 
client-server communication in their applications over 
a single persistent connection. 
Enabling more complex communication patterns caters for 
more interactive use cases, but introduces additional 
concerns to the developer with respect to implementation correctness.

@@

Whilst WebSockets make this web-based implementation possible, 
it introduces the developer to a new family of communication errors
(in addition to the usual testing for game logic correctness), 
even for this simple game:
we highlight just a few:

\begin{itemize}

\item
\textbf{Deadlocks:} how can we prevent both players waiting for 
each other to make a move at the same time?

\item 
\textbf{Communication mismatches:} what if player 1 sends
a boolean to the game server instead of board coordinates?

\item
\textbf{Channel linearity:} if the game server takes time
to update the game logic and respond to the players, what if 
player 1 clicks the same cell twice and sends the coordinates twice?

\end{itemize}

The complexity of these errors, which correlate to the complexity of tests 
required against these errors, scale with the complexity of the 
communication patterns involved. Over-reliance on integration testing
to attempt to expose communication-related bugs will also slow the
development process, not to mention that the time taken for these
integration tests would scale with the number of roles involved.
A localised, static way for verifying communication correctness would
be highly desirable.

@@

Existing work \cite{MVU2020,PureScript2019} that 
adapt the MPST framework for web development
acknowledge the limitations of JavaScript -- the language of the browser --
in providing static type-level guarantees, and proceed to use
languages equipped with stronger type systems that compile to JavaScript.
This comes at the cost of a learning curve for the developer,
and limits its utility in the mainstream web development space.

Among the many languages that compile to JavaScript,
TypeScript stands out as the most intuitive to use as it is defined
to be a \textit{superset} of JavaScript. It provides developers
with type-safety through its gradual, structural type system.
Whilst some \cite{MVU2020} point out this limits its usability for
encoding multiparty session types, we believe that the language
offers sufficient features that we can use to provide developers with
communication safety guarantees whilst preserving a flexible, natural
and idiomatic workflow. By building our work upon TypeScript, we
work towards incorporating MPST into mainstream web development,
which reduces development time by programmatically verifying 
implementations for communication correctness.
